State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-05-13 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex

%<-comment out this for checking equations (author list too)
\usepackage{iopams,setstack} 
%<-comment out this for checking equations
\usepackage{graphicx}
\usepackage{algorithm, algorithmicx, algpseudocode} 
\usepackage{color} 
%This is to decrease the space in subfigure command
\usepackage[tight,scriptsize]{subfigure}
\makeatletter
\renewcommand{\subfigtopskip}{1\p@}
\renewcommand{\subfigcapskip}{0\p@}
\renewcommand{\subfigcaptopadj}{1\p@}
\renewcommand{\subfigbottomskip}{1\p@}
\renewcommand{\subfiglabelskip}{0.05em plus 0.02em minus 0.01em}
\makeatother
% \usepackage{pstool}
\renewcommand{\subcapsize}{\small}
 \usepackage{anysize}							%<-use this for checking equation length
 % \marginsize {1.7cm}{1.5cm}{0.3cm}{0.8cm}			%<-use this for checking equation length
 %  \setlength{\columnsep}{0.5cm}					%<-use this for checking equation length
 % \usepackage{amsmath}
\usepackage{amssymb,amsfonts} 			%<-use this for checking equation length
\newcommand{\todo}[1]{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}} 
\newcommand{\wtf}[1]{\textsf{\emph{\textbf{\textcolor{red}{#1}}}}} 
\newcommand{\omg}[1]{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}} \begin{document}

@@

\address{$^1$ Department of Electrical and Electronic Engineering, University of Melbourne, Melbourne, Vic, 3010 Australia} \address{$^2$ Department of Automatic Control and Systems Engineering, University of Sheffield, Mappin Street, Sheffield, S1 3JD, UK} \address{$^3$ Department of Applied Physics and Applied Mathematics, Columbia University, US} \address{$^4$ Ken's} \ead{dfreestone@bionicear.org}

@@

\section{Neural Field Model}\label{NeuralModelSection} In this section, we describe a variant of the WC or Amari style neural field model \cite{Wilson1973,Amari1977} that was used to create our intracortical connectivity estimator. Neural field models relate mean firing rates of pre-synaptic neural populations to mean post-synaptic membrane potentials. They are popular due to being parsimonious yet having a strong link with the underlying physiology. Each neural population represents a functional cortical processing unit, such as a column. The columnar organisation of the cortex is continuous, where pyramidal cells are members of many columns. In general, cortical structure can be modelled in a physiologically plausible manner as being locally homogeneous (in short range intracortical connectivity) and heterogeneous (in long range cortico-cortical and corticothalamic connectivity)~\cite{Jirsa2009,Qubbaj2007}. In certain regoins of the cortex, each column is thought be connected locally via symmetric short range local excitation, with surround inhibition \cite{Braitenberg1998}. For example, this structural organisation is most studied in the visual system, where the surrounding inhibition effectively tunes a cortical column to a particular receptive visual field~\cite{Sullivan2006}. Neural field models are descriptive of a range of neurodynamics of the cortex such as evoked potentials, visual hallucinations, and epileptic behaviour~\cite{David2003,Bressloff2001,Breakspear2006}. 
Field models are also capable of generating complex patterns of activity such as Turing patterns, spirals, and traveling oscillations~\cite{Amari1977,Coombes2005,Coombes2007}.

\subsection{Derivation of the Integro-difference Equation Representation}
The model relates the average number of action potentials $g(\mathbf{r},t)$ arriving at position $\mathbf{r}$ to the local post synaptic membrane voltage $v(\mathbf{r},t)$. The post-synaptic potentials generated at a neuronal population at location $\mathbf{r}$ by action potentials arriving from all other connected populations at locations $\mathbf{r}'$ can be described by 
\begin{equation}
	\label{SpikesToPotential} v\left( {\mathbf{r},t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'}. 
\end{equation}
The post-synaptic response kernel $h(t)$ is described by 
\begin{equation}
	\label{SynapticRespKernel} h(t) = \eta(t)\exp{\left(-\zeta t\right)}. 
\end{equation}
where $\zeta=\tau^{-1}$ and $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function. Nonlocal interactions between cortical populations are described by 
\begin{equation}
	\label{RateBasedInteractions} g\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)d\mathbf{r}'}, 
\end{equation}
where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain, representing a cortical sheet or surface. The connectivity kernel is typically a ``Mexican hat'' function which describes strong local activation, weak mid-range repression and weak long-range activation. The exact shape of this kernel is assumed to vary across patients, and hence needs to be inferred from data.

@@

The firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function 
\begin{equation}
	\label{ActivationFunction} f\left( v\left( \mathbf{r}', t \right) \right) = \frac{f_{max}}{1 + \exp \left( \varsigma \left( v_0 - v\left(\mathbf{r}',t\right) \right) \right)}. 
\end{equation}
The parameters $f_{max}$ and $v_0$ describe the maximum firing rate and firing threshold of the neural populations. The parameter $\varsigma$ governs the slope of the sigmoid. By substituting equation~\ref{RateBasedInteractions} into \ref{SpikesToPotential} we get the spatiotemporal model 
\begin{eqnarray}
	\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) &=&  \int_{-\infty}^t h\left(t - t'\right) \\
	&\times&\int_\Omega w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}'dt'. \nonumber
\end{eqnarray}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function 
\begin{equation}
	\label{GreensFuncDef} Dh\left( t \right) = \delta \left( t \right), 
\end{equation}
where $D=\frac{d}{dt} + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{eqnarray}
	\label{FinalFormContinuous} \frac{dv\left( \mathbf{r},t \right)}{dt} &+& \zeta v\left( \mathbf{r},t \right) = \\
	&&\int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'}. \nonumber
\end{eqnarray}
To arrive at the integro-difference equation (IDE) form of the model we discretise time using a first-order Euler method (see~\ref{Time Discretization}) giving 
\begin{eqnarray}
	\label{DiscreteTimeModel} v_{t+T_s}\left(\mathbf{r}\right) &=& \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} \nonumber\\ 
	&+& e_t\left(\mathbf{r}\right), 
\end{eqnarray}
where $T_s$ is the time step, $\xi = 1-\zeta T_s$ and $e_t(\mathbf{r})$ is i.i.d. such that $e_t\sim\mathcal{GP}(\mathbf 0,\gamma(\mathbf{s}-\mathbf{r}))$. This term was added to account for uncertainty in the model. To generate iEEG/LFP data we used the output function 
\begin{equation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t, 
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field where $n=1,...,N$ indexes the sensors and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\Sigma_{\varepsilon}\right)$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the geometry of pick-up range of sensors where 
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)}. 
\end{equation}

@@

The spectral characteristics of the field are governed by the shape of the connectivity kernel and the covariance of the field disturbance. If the bandwidth of the field disturbance covariance is wider than the bandwidth of the connectivity kernel, then the disturbance covariance governs the bandwidth of the system. Conversely, a wider kernel bandwidth will govern the system bandwidth.

Let the spectral representation of the dynamic field at time $t$ be denoted by $V_t(\boldsymbol{\nu})$, where $\nu$ denotes spatial frequency. Given $V_t(\boldsymbol{\nu})$, the distance between adjacent sensors should be no more than
% 
% Then, for an accurate representation, $V_t(\boldsymbol{\nu})$ needs to be spatially band-limited. Such a condition is usually satisfied due to the band-limiting effect of most practical kernels and covariance functions \todo{~cite\{some EEG literature here\}} \omg{will try to find something. These kernel are kinda theoretical and I have not ever seen and estimate of the covariance function. Can we remove the sentence ``Such a condition is usually satisfied...'' }. Thus we can assume that the spectral response $V_t(\boldsymbol{\nu})$ is band-limited with $V_t(\boldsymbol{\nu}) \approx 0 ~ \forall \boldsymbol{\nu} > \boldsymbol{\nu}_c$,  
% 
% Note that for the two-dimensional, homogeneous and isotropic fields being considered $\boldsymbol{\nu}_c = [\nu_c ~ \nu_c]^\top$. Given such a band-limited field the distance between adjacent sensors, $\Delta_s$, needs to satisfy 
\begin{equation}
	\label{eq:MinimumSensorDistance} \Delta_s < \frac{1}{2\rho_s\nu_{c}}, 
\end{equation}
where $\boldsymbol{\nu}_c$ the cut-off frequency, $\rho_s \in \mathbb{R} \ge 1$ is an oversampling parameter. This condition must be satisfied for the hidden dynamic field $v_t(\mathbf{r})$ to be precisely inferred from the sampled observations, $\mathbf{y}_t$. If the condition is not satisfied, spatial aliasing will occur.

There are two options available to alleviate problems with aliasing. The first, which is the most desirable, is to increase the spatial sampling frequency (using a higher resolution of sensors). The second is to use wider sensors, where the spatial averaging acts like a low-pass anti-aliasing filter. This option is a compromise between alleviating issues with aliasing and losing high frequency information contained in the field. In practice, it is difficult  to estimate the bandwidth of the cortex using traditional electrophysiological measurements due to spatial limitations. However, we envisage it may be possible to estimate it using other modalities with higher spatial resolution such as, fMRI, NIRS or other optical imaging techniques~\cite{Issa2000}. Errors associated with mild aliasing result in distortions in the high frequency components of the field dynamics. However, after applying the basis function decomposition, high frequency components contribute negligibly to the inverse \omg{problem} and are therefore not a major problem for the state and parameter estimation~\cite{Sanner1992}. This is further illustrated our results. \omg{Although our results are severely effected by the loss of high frequency info! Ken: Do you think you can word this a little better than what I have done.}

Similar considerations need to be made regarding the representation of the dynamic field $v_t(\mathbf{r})$ using the basis function decomposition. The minimum distance between adjacent basis functions must also satisfy Shannon's sampling theorem to capture the dominant spectral information of the field. Thus, the minimum distance between basis functions must satisfy 
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_b < \frac{1}{2\rho_b\nu_{co}}. 
\end{equation}
where $\rho_b$ is an over-sampling parameter to determine the basis function separation and $\nu_{co}$ is the spatial cut-off frequency from the observations. For the Gaussian basis functions, $\phi(\mathbf{r})$, being considered, the basis function width parameter, $\sigma_{\phi}^2$, can also be inferred using spectral considerations~\cite{Sanner1992,Scerri2009}. The Fourier transform of an \textit{n}-dimensional Gaussian
\begin{equation}\label{eq:nDimensionalGaussian}
	\phi(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2} \mathbf r^\top\mathbf r}\right)
\end{equation}
 is another Gaussian given by
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
The basis function width, $\sigma^2_{\phi}$, should be chosen to give at least 3~dB attenuation at $\boldsymbol\nu_{co}$ giving \omg{need to rewrite eq~\ref{eq:WidthFrequencyRelationship} in terms of $\sigma_{\phi}$. At the mo it is writen in terms of $\sigma_{nu}$}
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\phi}= \frac{1}{\pi^2\sigma_{\nu_{co}}^2},
\end{equation}
where
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu_{co}}= \frac{2\boldsymbol\nu_{co}^\top \boldsymbol\nu_{co}}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent field with frequency content up to $\boldsymbol\nu_{co}$. Proofs for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in~\ref{ap:FrequencyAnalysis}.
% Thus to ensure that the basis functions can represent fields with frequency content up to $\boldsymbol{\nu}_c$, $\sigma_{\nu}$ should be set with at least the 3~dB frequency for the basis functions equal to $\boldsymbol{\nu}_c$, thus obtained $\sigma^2_{\nu}$ and therefore $\sigma_{\phi}^2$.
The dynamic field must be spatially oversampled using a larger number of sensors then imposed by frequency analysis considerations, to use frequency analysis to determine the minimum distance and width of the basis functions. Following this, the distance between basis functions is usually larger then the distance between sensors, although this is not a requirement for using this estimation framework. \omg{Ken: can you please write something about what should be done in this case?} For a spatially homogeneous isotropic fields, the basis functions can be placed on a regular grid. Following this, the knowledge of the distance between basis functions directly implies the total number of basis functions required.

@@

POSSIBLE SOURCES OF THE ERROR
\begin{itemize}
	\item slope of nonlinearity interacting with tails of Gaussian basis functions
	\item loss of high frequency info from basis function decomposition
	\item distortions of the spectral content of the basis function from the nonlinearity
	\item disturbance covariance to narrow
	\item disturbance variance to low (no exciting the kernel enough)
	\item boundary conditions
	\item
\end{itemize}

>>>>

%<-comment out this for checking equations (author list too)
\usepackage{iopams,setstack} 
%<-comment out this for checking equations
\usepackage{graphicx}
\usepackage{algorithm, algorithmicx, algpseudocode} 
\usepackage{color} 
%This is to decrease the space in subfigure command
\usepackage[tight,scriptsize]{subfigure}
\makeatletter
\renewcommand{\subfigtopskip}{1\p@}
\renewcommand{\subfigcapskip}{0\p@}
\renewcommand{\subfigcaptopadj}{1\p@}
\renewcommand{\subfigbottomskip}{1\p@}
\renewcommand{\subfiglabelskip}{0.05em plus 0.02em minus 0.01em}
\makeatother
% \usepackage{pstool}
\renewcommand{\subcapsize}{\small}
 \usepackage{anysize}							%<-use this for checking equation length
 % \marginsize {1.7cm}{1.5cm}{0.3cm}{0.8cm}			%<-use this for checking equation length
 %  \setlength{\columnsep}{0.5cm}					%<-use this for checking equation length
 % \usepackage{amsmath}
\usepackage{amssymb,amsfonts} 			%<-use this for checking equation length
\newcommand{\todo}[1]{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}} 
\newcommand{\wtf}[1]{\textsf{\emph{\textbf{\textcolor{red}{#1}}}}} 
\newcommand{\omg}[1]{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}}
\newcommand{\haq}[1]{\textsf{\emph{\textbf{\textcolor{magenta}{#1}}}}}
\begin{document}

@@

\address{$^1$ Department of Electrical and Electronic Engineering, University of Melbourne, Melbourne, Vic, 3010 Australia} \address{$^2$ Department of Automatic Control and Systems Engineering, University of Sheffield, Mappin Street, Sheffield, S1 3JD, UK} \address{$^3$ Department of Applied Physics and Applied Mathematics, Columbia University, US} \address{$^4$ Deaprtment of Systems and Control Engeineering, University of Malta, Msida, MSD 1333, Malta} \ead{dfreestone@bionicear.org}

@@

\section{Neural Field Model}\label{NeuralModelSection} In this section, we describe a variant of the WC or Amari style neural field model \cite{Wilson1973,Amari1977} that was used to create our intracortical connectivity estimator. Neural field models relate mean firing rates of pre-synaptic neural populations to mean post-synaptic membrane potentials. They are popular due to being parsimonious yet having a strong link with the underlying physiology. Each neural population represents a functional cortical processing unit, such as a column. The columnar organisation of the cortex is continuous, where pyramidal cells are members of many columns. In general, cortical structure can be modelled in a physiologically plausible manner as being locally homogeneous (in short range intracortical connectivity) and heterogeneous (in long range cortico-cortical and corticothalamic connectivity)~\cite{Jirsa2009,Qubbaj2007}. In certain regions of the cortex, each column is thought to be connected locally via symmetric short range local excitation, with surround inhibition \cite{Braitenberg1998}. For example, this structural organisation is most studied in the visual system, where the surrounding inhibition effectively tunes a cortical column to a particular receptive visual field~\cite{Sullivan2006}. Neural field models are descriptive of a range of neurodynamics of the cortex such as evoked potentials, visual hallucinations, and epileptic behaviour~\cite{David2003,Bressloff2001,Breakspear2006}. 
Field models are also capable of generating complex patterns of activity such as Turing patterns, spirals, and traveling oscillations~\cite{Amari1977,Coombes2005,Coombes2007}.

\subsection{Derivation of the Integro-difference Equation Representation}
The model relates the average number of action potentials $g(\mathbf{r},t)$ arriving at position $\mathbf{r}$ to the local post synaptic membrane voltage $v(\mathbf{r},t)$. The post-synaptic potentials generated at a neuronal population at location $\mathbf{r}$ by action potentials arriving from all other connected populations at locations $\mathbf{r}'$ can be described by 
\begin{equation}
	\label{SpikesToPotential} v\left( {\mathbf{r},t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'}. 
\end{equation}
The post-synaptic response kernel $h(t)$ is described by 
\begin{equation}
	\label{SynapticRespKernel} h(t) = \eta(t)\exp{\left(-\zeta t\right)}. 
\end{equation}
where $\zeta=\tau^{-1}$ and $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function. Nonlocal interactions between cortical populations are described by 
\begin{equation}
	\label{RateBasedInteractions} g\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)d\mathbf{r}'}, 
\end{equation}
where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain, representing a cortical sheet or surface. The connectivity kernel is typically a ``Mexican hat'' function which describes strong local activation, weak mid-range repression and weak long-range activation. The exact shape of this kernel is assumed to vary across patients, and hence needs to be inferred from data. \haq{Risky comment if we don't get better estimates.}

@@

The firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function 
\begin{equation}
	\label{ActivationFunction} f\left( v\left( \mathbf{r}', t \right) \right) = \frac{f_{max}\left(\mathbf{r}\right)}{1 + \exp \left( \varsigma \left( v_0\left( \mathbf{r} \right) - v\left(\mathbf{r}',t\right) \right) \right)}. 
\end{equation}
The parameters $f_{max}$ and $v_0$ describe the maximum firing rate and firing threshold of the neural populations. The parameter $\varsigma$ governs the slope of the sigmoid. By substituting equation~\ref{RateBasedInteractions} into \ref{SpikesToPotential} we get the spatiotemporal model 
\begin{eqnarray}
	\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) &=&  \int_{-\infty}^t h\left(t - t'\right) \\
	&\times&\int_\Omega w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}'dt'. \nonumber
\end{eqnarray}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function 
\begin{equation}
	\label{GreensFuncDef} Dh\left( t \right) = \delta \left( t \right), 
\end{equation}
where $D=\frac{d}{dt} + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{eqnarray}
	\label{FinalFormContinuous} \frac{dv\left( \mathbf{r},t \right)}{dt} &+& \zeta v\left( \mathbf{r},t \right) = \\
	&&\int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'}. \nonumber
\end{eqnarray}
To arrive at the integro-difference equation (IDE) form of the model we discretise time using a first-order Euler method (see~\ref{Time Discretization}) giving 
\begin{eqnarray}
	\label{DiscreteTimeModel} v_{t+1}\left(\mathbf{r}\right) &=& \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} \nonumber\\ 
	&+& e_t\left(\mathbf{r}\right), 
\end{eqnarray}
where $T_s$ is the time step, $\xi = 1-\zeta T_s$ and $e_t(\mathbf{r})$ is i.i.d. such that $e_t\sim\mathcal{GP}(\mathbf 0,\gamma(\mathbf{s}-\mathbf{r}))$. \haq{Can we remove noise form 8 and introduce only discretization there; then introduce noise in another equation.} This term was added to account for uncertainty in the model. To represent the iEEG/LFP data collection precedure we used the output function 
\begin{equation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t, 
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field where $n=1,...,N$ indexes the sensors and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\Sigma_{\varepsilon}\right)$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the geometry of pick-up range of sensors where 
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)}. 
\end{equation}

@@

Let the spectral representation of the dynamic field at time $t$ be denoted by $V_t(\boldsymbol{\nu})$. Then, based on Shannon's sampling Theorem, for an accurate representation of this field by spatially discrete observations, $V_t(\boldsymbol{\nu})$ needs to be spatially band-limited. Such a condition ensures that $V_t(\boldsymbol{\nu}) \approx 0 ~ \forall \boldsymbol{\nu} > \boldsymbol{\nu}_c$, where $\boldsymbol{\nu}_c$ is a cutoff frequency with  $\boldsymbol{\nu}_c = [\nu_c ~ \nu_c]^\top$. Given such a band-limited field, the distance between adjacent sensors, $\Delta_s$, needs to satisfy 
\begin{equation}
	\label{eq:MinimumSensorDistance} \Delta_s < \frac{1}{2\rho_s\nu_{c}}, 
\end{equation}
where $\rho_s \in \mathbb{R} \ge 1$ is an oversampling parameter. This condition must be satisfied to avoid spectral aliasing effects in the representation of the hidden dynamic field $v_t(\mathbf{r})$ by the sampled observations, $\mathbf{y}_t$.

In practice, it is difficult to estimate the bandwidth of the cortex using traditional electrophysiological measurements thus allowing the positioning of sensor in accordance to \ref{eq:MinimumSensorDistance}. However, we envisage it may be possible to estimate it using other modalities with higher spatial resolution such as, fMRI, NIRS or other optical imaging techniques~\cite{Issa2000}. Nevertheless, using electrophysiological measurements, spectral aliasing can still be avoided by a proper choice of the spatial sampling distance given the sensor's spectral characteristics. Sensors with smooth wide spatial responses result in a spectral low-pass action, thus providing spatial anti-alaising filtering. Such sensors therefore bandlimit the field to a known bandwidth, denoted by $\nu_{co}$. Aliasing can then be avoided by positioning the sensors according to \ref{eq:MinimumSensorDistance}, with $\nu_{co}$ replacing $\nu_c$.

Although such sensors avoid errors due to alaising, they attentuate the high frequency variations in the observations. Therefore, any procedure applied to estimate the original field or the kernel from these bandlimited observations will underestimate the high frequency component in both the field and the kernel. Such a remark motivates the use of sensors with wider bandwidths. Nevertheless, this choice would require an excessive number or sensors in order to satisfy Shannon's sampling theorem for a reasonable spatial region. Therefore, a compromise needs to be found between the bandwidth of the sensor response and therefore the accurate of the estimation results and the number of sensors used and thus the computational demands of estimation procedure.

Similar considerations need to be made regarding the representation of the dynamic field $v_t(\mathbf{r})$ using the basis function decomposition. The minimum distance between adjacent basis functions must also satisfy Shannon's sampling theorem to capture the dominant spectral information of the field. Thus, the minimum distance between basis functions must satisfy 
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_b < \frac{1}{2\rho_b\nu_{co}}. 
\end{equation}
where $\rho_b$ is an over-sampling parameter to determine the basis function separation. For the Gaussian basis functions, $\phi(\mathbf{r})$, being considered, the basis function width parameter, $\sigma_{\phi}^2$, can also be inferred using spectral considerations~\cite{Sanner1992,Scerri2009}. The Fourier transform of an \textit{n}-dimensional Gaussian is another Gaussian given by

\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
The basis function width, $\sigma^2_{\phi}$, should be chosen to give at least 3~dB attenuation at $\boldsymbol\nu_{co}$ giving \omg{need to rewrite eq~\ref{eq:WidthFrequencyRelationship} in terms of $\sigma_{\phi}$. At the mo it is writen in terms of $\sigma_{nu}$}
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\phi}= \frac{1}{\pi^2\sigma_{\nu_{co}}^2},
\end{equation}
where
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu_{co}}= \frac{2\boldsymbol\nu_{co}^\top \boldsymbol\nu_{co}}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent field with frequency content up to $\boldsymbol\nu_{co}$. Proofs for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in~\ref{ap:FrequencyAnalysis}.

Note that for a spatially homogeneous isotropic fields, the basis functions can be placed on a regular grid. Thus, the knowledge of the distance between basis functions directly implies the total number of basis functions required to represent a known spatial region.

@@
